---
title: "Lab 8"
author: "Vaughan Andrews"
date: "3/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, include=FALSE}

library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)
library(corrplot)

```

```{r load_data}

oil_spills <- read.csv("oil_spills.csv")

```

#### Column graph of Texas oil spills

```{r}

df <- oil_spills %>% 
  filter(`Accident.State` =="TX" & `Accident.Year` < 2017) %>% 
  group_by(`Accident.Year`) %>% 
  summarise(Loss = sum(`Net.Loss..Barrels.`))

colnames(df) <- c("Year","Loss")

ggplot(df, aes(x = Year, y = Loss))+
  geom_col() +
  theme_classic()
         

```

#### Leaflet plot of spill locations in TX in 2016

```{r}

df_loc <- oil_spills %>% 
  filter(`Accident.State` == "TX", `Accident.Year` == 2016) %>% 
  select(Latitude,Longitude, `Net.Loss..Barrels.`)

colnames(df_loc) <- c("latitude","longitude","net_loss")

oil_sf <- st_as_sf(df_loc, coords = c("longitude","latitude"), crs = 4326) #convert to simple features for use in leaflet
 

leaflet(oil_sf) %>% 
  addTiles() %>% 
  addMarkers()

```

#### tmap plot with the Texas state shapefile

```{r}

states <- st_read(dsn = ".", layer = "states")

tex_border <- states %>% 
  filter(STATE_NAME == "Texas") %>%
  st_transform(4326)

plot(tex_border)

tm_shape(tex_border) +
  tm_polygons() +
  tm_shape(oil_sf) +
  tm_dots(size = 0.3)

```

#### Convert the data to spatial point patterns (combination of point data and the bounding window)

```{r}

spill_sp <- as(oil_sf, "Spatial")
spill_ppp <- as(spill_sp, "ppp") #convert to point pattern for point pattern analysis

tx_sp <- as(tex_border, "Spatial")
tx_owin <- as(tx_sp, "owin")

all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin)

plot(density(all_ppp, sigma = 0.3)) #use with caution

```

#### Quadrat test for spatial evenness

```{r}

oil_qt <- quadrat.test(all_ppp, nx = 5, ny = 5) #5 regions horizontally and vertically
oil_qt #null hypothesis is complete spatial randomness (CSR), conlude that the data are NOT evenly distributed

plot(all_ppp)
plot(oil_qt, add = TRUE, cex = 0.4) #plots the quadrats (shows actual counts, expected counts, std measure of difference)

```

#### G-function for Nearest Neighbor Analysis

```{r}

r <- seq(0,1, by = 0.01) #choose this range to start with but you'll have to change depending on data

oil_gfun <- envelope(all_ppp, fun = Gest, r = r, nsim = 100) #the entire space dataset, using the G-estimates for G-function, 100 simulations

ggplot(oil_gfun, aes(x = r, y = obs)) + #our actual data
  geom_line(color = "black") +
  geom_line(aes(x = r, y = theo), color = "red") + #modeled CSR data, theo comes from the envelope function
  theme_classic()
```

#### Nearest neighbor using the L-function (Ripley's K, standardized), more advanced analysis than G (radius vs line for NN)

```{r}

r2 <- seq(0,5, by = 0.5) #different range because allison played around first

oil_lfun <- envelope(all_ppp, fun = Lest, r = r2, nsim = 10, global = TRUE) #include global to include all observations

ggplot(oil_lfun, aes(x = r2, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x = r2, y = theo), color = "blue") +
  theme_classic()

```

Four different ways to look at spatial distribution.
1. Look at data
2. Quadrat analysis
3. G function
4. L function